---
title: "Trees Compaison Siciliano"
output: html_notebook
---

Import libraries and attach data
```{r}
library(ISLR2)
library(tree)
#data(Carseats)
#attach(Carseats)

carseats_train = read.csv("Carseats_train.csv")

carseats_test = read.csv("Carseats_test.csv")


```

Create classification variable , required for regression

```{r}

suppressMessages(attach(carseats_train))


carseats_train = subset(carseats_train, select = -X ) #drops x column

#High_train <- factor(ifelse(Sales < 8, "No", "Yes")) #creates high variable
#carseats_train <- data.frame(carseats_train , High_train) # attaches high variable

#View(carseats_train)

suppressMessages(attach(carseats_test))

carseats_test = subset(carseats_test, select = -X ) #drops x column


#High_test <- factor(ifelse(Sales < 8, "No", "Yes")) #creates high variable
#carseats_test <- data.frame(carseats_test , High_test) # attaches high variable

#View(carseats_test)



```

required variables for a classification tree 

```{r}

suppressMessages(attach(carseats_train))


carseats_train = subset(carseats_train, select = -X )

High_train <- factor(ifelse(Sales < 8, 0, 1)) #creates high variable  did have factor() here 
carseats_train <- data.frame(carseats_train , High_train) # attaches high variable



#class(carseats_train$High_train)
#carseats_train$High_train = as.integer(carseats_train$High_train)
#class(carseats_train$High_train)



suppressMessages(attach(carseats_test))

carseats_test = subset(carseats_test, select = -X )


High_test <- factor(ifelse(Sales < 8, 0, 1)) #creates high variable
carseats_test <- data.frame(carseats_test , High_test) # attaches high variable



```

```{r}
carseats_train$ShelveLoc = as.factor(carseats_train$ShelveLoc)
carseats_train$US = as.factor(carseats_train$US)
carseats_train$Urban = as.factor(carseats_train$Urban)


```



```{r}
head(carseats_train)
```



Classification tree 

```{r}
suppressMessages(attach(carseats_train))


#View(Carseats)


#classification
tree.carseats <- tree(High_train ~ . - Sales, carseats_train, split= "gini") #dropping sales here 



#regression
#tree.carseats <- tree(Price ~ . - Price, carseats_train) #dropping sales here 
summary(tree.carseats)


```
```{r}

plot(tree.carseats)
text(tree.carseats, pretty = 0)



```



```{r}
View(tree.carseats$frame)

```



Prediction 

```{r}
#attach(carseats_test)

tree.pred <- predict(tree.carseats , carseats_test ) #,type = "tree"
#View(tree.pred)
#View(Price)

mse = 0 

#print(typeof(tree.pred))
#print(length(tree.pred))

for(i in 1:length(tree.pred))
  
  mse = mse + (Price[i] - tree.pred[i])**2
  #print(mse)
print(mse/length(tree.pred))

#MSE 378.2664 full tree

```

Pruning with Misclassification

```{r}
#attach(carseats_train) # changeable with test and train depending on which graph wanting results 


prune.carseats <- prune.tree(tree.carseats, best =6)  #in order to prune the tree (9 refers to the "size")

#sequences of cuts, leaves:misclass :  13:378.2664,  12:374.0569,  11:372.0732,   10:379.3271,   8: 401.6294 , 7:377.1002, 6: 365.1234, 5:377.2487,  4:403.2813, 3:440.0371,    2: 458.5094#Test
# 13:191.7746,  12:199.1575,  11:206.63 ,   10:214.9282,   8:244.6949 , 7:261.4313, 6: 281.0706 , 5:301.3909,  4:332.2309, 3:386.2193,    2:443.0949  #train

#plot(prune.carseats)
#text(prune.carseats , pretty = 0)
#summary(prune.carseats)
#prune.carseats

tree.pred <- predict(prune.carseats , carseats_test ) #,type = "class")


#attach(carseats_test) #carseats_test


mse = 0 
for(i in 1:length(tree.pred))
  mse = mse + (Price[i] - tree.pred[i])**2
print(mse/length(tree.pred))


```

```{r}

plot(prune.carseats)
text(prune.carseats, pretty = 0)
```


Research

Does the tree libarary in R grow with gini or entropy 

when using tree there are only the split criteria as "deviance" or "gini"












another attempt using gini

```{r}
set.seed(1)
n=200
X1=runif(n)
X2=runif(n)
P=.8*(X1<.3)*(X2<.5)+
  .2*(X1<.3)*(X2>.5)+
  .8*(X1>.3)*(X1<.85)*(X2<.3)+
  .2*(X1>.3)*(X1<.85)*(X2>.3)+
  .8*(X1>.85)*(X2<.7)+
  .2*(X1>.85)*(X2>.7) 
Y=rbinom(n,size=1,P)  
B=data.frame(Y,X1,X2)
```



```{r}
tree(Y~X2,method="gini")
```

