---
title: "Thesis_best_subset"
author: "Daniel Fynn"
date: "2023-08-12"
output: html_document
---


fit separate least squared models to the data

choice dataset: carseats, price as predictor, transforming categorical predictors to dummy encoding

Following introduction to statistical learning Chapter 6 Lab, 220


```{r}
carseats_train = read.csv("Carseats_train.csv")


carseats_train = subset(carseats_train, select = -X )

carseats_train

```

regsubsets function() from leaps library


```{r}

library(leaps)

regfit.full = regsubsets(Price ~.,carseats_train, nvmax =11 )
summary(regfit.full)

```
Asterixes show variable inclusion for different amount of predictors giving best response for lm. 
Showing comprice as the giving best prediction for a single variable. 


R^2 statistic

```{r}
summary(regfit.full)$rsq
```

rss vs adjusted R^2

```{r}

par(mfrow =c(2,2))
plot(summary(regfit.full)$rss ,xlab=" Number of Variables ",ylab=" RSS", type="l")
plot(summary(regfit.full)$adjr2 ,xlab =" Number of Variables ",
ylab=" Adjusted RSq",type="l")


points (9, summary(regfit.full)$adjr2[9], col ="red",cex =2, pch =20)


plot(summary(regfit.full)$cp ,xlab =" Number of Variables ",ylab="Cp", type="l")

points(8, summary(regfit.full)$cp [8], col ="red",cex =2, pch =20)

plot(summary(regfit.full)$bic ,xlab=" Number of Variables ",ylab=" BIC", type="l")
points(7, summary(regfit.full)$bic [7], col =" red",cex =2, pch =20)

```

```{r}
which.max (summary(regfit.full)$adjr2)
which.min (summary(regfit.full)$cp )
which.min (summary(regfit.full)$bic )
```

```{r}
plot(regfit.full ,scale ="r2")
```

```{r}
plot(regfit.full ,scale ="adjr2")

```


```{r}
plot(regfit.full ,scale ="Cp")
```

```{r}
plot(regfit.full ,scale ="bic")
```


```{r}
coef(regfit.full ,7)
```


cross validation 


```{r}
carseats_test = read.csv("Carseats_test.csv")
carseats_test = subset(carseats_test, select = -X )


```



```{r}
test.mat=model.matrix(Price ~., data=carseats_test)

```



```{r}
val.errors =rep(NA ,11)

for(i in 1:11){
 coefi=coef(regfit.full ,id=i)
 pred=test.mat[,names(coefi)]%*% coefi
 val.errors[i]= mean((carseats_test$Price-pred)^2)
}
 
```


```{r}
val.errors
```

```{r}
which.min (val.errors )

coef(regfit.full ,7)
```

predict fn

```{r}
predict.regsubsets = function (object ,newdata ,id ,...){
 form=as.formula(object$call [[2]])
 mat=model.matrix(form ,newdata )
 coefi =coef(object ,id=id)
 xvars =names (coefi )
 mat[,xvars ]%*% coefi
 }
```




find the best 7 variable model of the whole dataset, not just the training set


```{r}
library("ISLR")
data(Carseats)

```



```{r}
regfit.best=regsubsets(Price ~.,data=Carseats ,nvmax =11)
coef(regfit.best ,7)
```


10 k folds


```{r}

k = 10
set.seed(1)
folds=sample(1:k,nrow(Carseats ),replace =TRUE)
cv.errors =matrix(NA ,k,11, dimnames =list(NULL , paste (1:11) ))


```




```{r}

for(j in 1:k){
 best.fit =regsubsets(Price ~.,data=Carseats[folds !=j,], nvmax =11)
 for(i in 1:11) {
 pred=predict.regsubsets(best.fit ,Carseats[folds ==j,], id=i)
 cv.errors[j,i] = mean( (Carseats$Price[folds ==j] - pred)^2)
 }
 }

```



```{r}

mean.cv.errors =apply(cv.errors ,2, mean)
mean.cv.errors
plot(mean.cv.errors ,type="b")
which.min(mean.cv.errors)
```


```{r}
reg.best =regsubsets(Price ~.,data=Carseats, nvmax =11)
coef(reg.best ,7)
```



```{r}
data(Carseats)

Carseats = subset(Carseats, select = -Population)
Carseats = subset(Carseats, select = -Education)
Carseats = subset(Carseats, select = -Urban)
Carseats = subset(Carseats, select = -US)

lm1 = lm(Price ~.,data=Carseats)  
summary(lm1)
```





